Testing Quasar Unification with Clumpy Wind Models

### 2015-11-10 paper.tex
bringing up to date with polarisation etc

allows for a more realistic AGN luminosities,
while maintaining the ionization state necessary for strong BAL features in the rest-frame UV. 

@@

The dense, X-ray heated wind 
produces strong recombination and collisionally excited line emission in, 
e.g., \civ\ and \la, to emerge at the low inclination, `Type 1 quasar-like' angles. 
At the highest inclinations, the synthetic spectra possess prominent \mg\ and \al\ BALs, the absorption 
features seen in LoBAL quasars. 

@@

{\bf This sentence needs to be more precise, especially in 
light of the statement earlier in the abstract about ``good agreement''}

@@

{\bf There is confusion in the abstract, at least, in what is the model and what is the synthetic spectra, as well as whether there is one model or many models.  We need to try to be clear.  The model should be a clumped biconical outflow.  There are then particular instantiations of the model. And then there are synthetic spectra}

@@

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is often attributed to emssion from a 
geometrically thin, optically thick accretion disc surrounding the central black hole, similar
to that described by \cite{shakurasunyaev1973}.  The lines are usually attributed to scattering and absorption of material above the disk.  
{\bf Note: the first half of this paragraph and the second do not follow from one another.  
The first half is about spectra the second about outflows and the connection to spectra is not made until the next paragrph.  
This needs fixing. JM: Knox, I like this structure...I start with inflow, then move to outflow.}
In addition to the inflowing accreting material, 
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution \citep{springel2005}
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

Approximately $20\%$ of quasars exhibit blueshifted ($\sim 0.1c$)
broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within the ADW paradigm, a biconical wind rises from 
the accretion disc and the BALQSO fraction is associated with
the covering factor of the outflow. 
Polarisation studies expect the wind to be roughly equatorial
\citep{goodrich1995, cohen1995}, although this geometry is contradicted
by more recent studies of radio spectral indices (REFS DIPOMPEO; see section ?).
ADWs offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) can 
correspond either to the dense wind base or dense clumps embedded
in the outflows. Indeed, \cite{elitzur2014} show that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit high ionization line absorption; LoBALQSOs, which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified geometric models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

% As well as acting as a source of photoionized plasma, a 
As well as imprinting clear line absorption and emission
features, disc winds may also have a profound effect on the structure and 
emergent {\em continuum} of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features typically seen in luminous AGN \citep{knigge1999,laordavis2014}.
There have been numerous difficulties when confronting 
theoretical accretion disc models with observations 
\citep[see e.g.][]{blaes1998}. However, AGN spectral energy distributions (SEDs) can now, 
in general, be fitted well with accretion disc models
when the effects of general relativity (GR), Comptonisation
and mass-loss are included \citep{capellupo2015}. Mass-loss therefore appears to be 
critical if an accretion disc model is to successfully fit AGN SEDs, 
particular in the UV region of the spectrum.

{\bf ksl: I don't see what this paragraph has to do with our problem, which we need to clearly explain} Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics, 
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms for ADWs have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
The presence of line-locked features \citep{bowler2014} 
and the `ghost of \la' (Arav et al. 1996; Arav 1996; North 2006; but see 
also Cottis et al. 2010) \nocite{arav1995, arav1996, north2006,cottis2010}
in the spectra of some BALQSOs also gives clearer evidence that line-driving is
at least partially contributing to the acceleration of the wind.

The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma, meaning that it is difficult to prevent 
the wind becoming over-ionized and `failing' in the presence of strong X-rays. 
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.
An additional or alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \cite{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required to explain the electron scattering wings of emission lines formed
in line-driven hot star winds \citep{hillier1991eswingsmodel}. Complex substructures 
on a variety of scales are also produced in simulations of line-driven 
outflows in AGN \citep{PSK2000,PK04,progakurosawa2010,proga2014}.
Clumpy winds therefore offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

@@

Our approach to line transfer is based upon the macro-atom implementation developed by \cite{lucy2002, lucy2003}. In our case, for reasons of computational effieciency, we adopt the  hybrid macro-atom scheme 
described by M15 (an improvement over that used by H14 for our earlier quasar study), 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interact with either two-level `simple ions'
or full `macro-atoms'. 
Our scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H is treated as a macro-atom, because 
we expect recombination to be important
in determining their level populations and resultant line emission, and because we were especially interested in the contribution to AGN spectra of \LA.  In H14, all atoms were treated in a two-level approximation.

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by M15. Previously (LK02, H13, M15),
we used a modified Saha approach to calculate the ionization fractions
of simple-ions. As part of  this effort, we have now improved {\sc PYTHON} to explicitly solve the 
rate equations between ions in non-LTE. 
This dispenses with a number of small assumptions made in the modified Saha approach, 
is more numerically stable, 
and, in principle, allows the direct addition of extra physical 
processes such as Auger ionization.

{\bf ksl - Do we need the rest of this, or could it be postponed until a detailed paper on the program }

%To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
We model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity in a series of $n$ bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $i$, or an exponential 
\begin{equation}
J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{DR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination, respectively.
%% We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

% We improve on this method by abandoning the modified Saha approach
% entirely, and instead computing the ionization state by 
% explicitly solving the rate equations in between ions in non-LTE. 
% Photoionization and recombination rates are calculated using MC estimators recorded
% during the photon propagation. Not only does this dispense with a number of
% small assumptions made in the modified Saha approach, it is also more numerically stable, 
% and in principle allows the direct addition of extra physical processes such as Auger ionization, which would have to be approximated if using the previous technique.
\subsection{Physical Processes}

@@

% MODEL DESCRIPTION

@@

Our kinematic prescription for a biconical disc wind model
follows \cite{SV93}, and is described further by
LK02, H13 and M15. A schematic is shown in figure~\ref{fig:cartoon},
with key aspects marked. The general biconical
geometry is similar to that invoked by \cite{MCGV95} and 
\cite{elvis2000} in order to explain the phenomenonology
of quasars and BALQSOs.

@@

% \subsection{A Benchmark Model for BALQSOs}

% \cite{higginbottom2013} presented a benchmark model for
% (BAL)QSOs...introduce key parameters.

@@

%Describe the photon sources in the model.

The accretion disc in our model is geometrically thin, but optically thick.
We thus treat the disc as an ensemble of blackbodies with a 
\cite{shakurasunyaev1973} effective temperature profile. 
The emergent SED is then determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are assumed to beopaque, meaning
that photons which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~r_G$, where 
$r_G = GM_{BH}/c^2$ is the gravitational radius.
For a $10^9~M_\odot$ black hole, this is equal to $8.8\times10^{14}~{\rm cm}$ 
or $\sim10^{-4}~{\rm pc}$.

The X-ray source in our model is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet
constraint (see Lucy 2002, M15).
% \begin{equation}

% \end{equation}

% With a normalisation such that it produces the correct 2-10kev luminosity.

% Photons striking the

% which successfully reproduced BAL features 
% in synthetic spectra using \py. However, the model
% had a few key drawbacks when considering its potential
% for unification. First, it failed to produce significant
% broad emission lines at non-BAL viewing angles; this is 
% a key requirement for a unified QSO model. Second,
% the model was significantly X-ray weak, as it was found
% that an X-ray luminosity comparable to that of a QSO 
% over-ionized the wind, resulting in no UV resonance 
% line absorption features.

@@

In our model, a biconical disc wind rises from the accretion 
disc between launch radii $r_{min}$ and $r_{max}$.
The opening angles of the wind are set to $\theta_{min}$ and $\theta_{max}$.
The poloidal velocity along each individual streamline at a poloidal distance $l$ 
is then given by
\begin{equation}
v_l=v_0+\left[v_{\infty}(r_0)-v_0\right]\frac{\left(l/R_v\right)^{\alpha}}{\left(l/R_v\right)^{\alpha}+1},
\label{v_law}
\end{equation}
where $v_0$ is the velocity at the base of the streamline, $\alpha$ is
an exponent governing how quickly the wind accelerations and 
$R_v$ is the `acceleration length', defined as the distance at which
the outflow reaches half of its terminal velocity, $v_{\infty}$.
The terminal velocity is set to a fixed multiple of the escape
velocity, $v_{esc}$, at the base of the streamline (radius $r_0$).
The rotational velocity, $v_{\phi}$, is initially Keplerian ($v_k = [GM/r_0]^{1/2}$),
and the wind conserves specific angular momentum, such that 
\begin{equation}
v_{\phi} r = v_k r_0.
\label{v_law}
\end{equation}
The velocity law is crucial in determining the output spectra,
as it affects not only the projected velocities along the line of sight,
but also the density and ionization state of the outflow.
A wind which accelerates more slowly will have a denser wind base
with correspondingly different ionization and emission characteristics.

@@

% {\bf This needs an introductory paragraph.  Should state our earlier modeling of AGN H13 was based on a smooth outflow.  It should reference the introduction where we say there are observational indications of clumping.  Should say that the difficulty associated with clumping is two fold - computationally implies many grid cells,  many potential parameters.  Ultimately, say that for simplicity we use microclumping.  Then say this has been used elsewhere, but with more recent references than what we have.  The point is we need to motivate micro clumping before we say what it means to the code  It's all mixed up as written. Aside: at some point this paper should mention H14, and hydro models more generally}

To allow for clumping in our outflow we adopt a simple approximation
used extensively in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998}(MORE REFS). The key assumption here is that 
typical clump sizes are much smaller than the typical photon mean free path, 
and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `filling factor', $f$, which is the 
fraction of the volume of the plasma filled by clumps. The intra-clump
medium is assumed to be a vacuum. We can then introduce the 
density enhancement, $D$, which is defined as 
\begin{equation}
D = \frac{1}{f}.
\end{equation}
We then multiply all densities in the model by $D$, and all emitting volumes
by $f$. This has the effect of enhancing all emissivities and opacities
that scale with the square of density (such as collisional excitation and recombination) 
by a factor $D$. All processes that scale linearly with density 
(such as electron scattering and bound-free opacity)
will remain unchanged for a given ionization state. 
% A factor $f$ is also
% applied to the opacities such that opacities which scale only with $\rho$ are not
% increased.

@@

@@

{\bf We need to describe what we did before getting to results.  We started with parameters similar to those of H14, we created a small grid.  Limits of the grid are given in table 1.  This should be a paragraph or section before this one.  Not all quasars have to have the same parameters and so we have to explain that we have elected to describe one model in detail.  Was our goal to fix the luminosity and see what was required.

>>>>

allows for a more realistic AGN luminosities, while maintaining the ionization 
state necessary for strong BAL features in the rest-frame UV. 

@@

The dense, X-ray heated wind produces strong recombination and collisionally 
excited line emission in, e.g., \civ\ and \la, to emerge at the low inclination, 
`Type 1 quasar-like' angles. 
At the highest inclinations, the synthetic spectra possess prominent \mg\ and \al\ BALs, the absorption features seen in LoBAL quasars. 

@@

% {\bf This sentence needs to be more precise, especially in 
% light of the statement earlier in the abstract about ``good agreement''}

@@

% {\bf There is confusion in the abstract, at least, in what is the model and what is the synthetic spectra, as well as whether there is one model or many models.  We need to try to be clear.  The model should be a clumped biconical outflow.  There are then particular instantiations of the model. And then there are synthetic spectra}

@@

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is often attributed to emission from a 
geometrically thin, optically thick accretion disc surrounding the central black hole, similar
to that described by \cite{shakurasunyaev1973}.
% {\bf Note: the first half of this paragraph and the second do not follow from one another.  
% The first half is about spectra the second about outflows and the connection to spectra is not made until the next paragrph.  
% This needs fixing. JM: Knox, I like this structure...I start with inflow, then move to outflow.}
In addition to the {\em inflowing} accreting material, 
{\em outflows} are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution \citep{springel2005}
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

Perhaps the clearest evidence of outflows in AGN is  
the blueshifted ($\sim 0.1c$) broad absorption lines (BALs) in the 
ultraviolet seen in approximately $20\%$ of quasars
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within the ADW paradigm, a biconical wind rises from 
the accretion disc and the BALQSO fraction is associated with
the covering factor of the outflow. 
Polarisation studies expect the wind to be roughly equatorial
\citep{goodrich1995, cohen1995}, although there is also evidence
for polar BAL outflows \citep{zhou2006}.
% although this geometry is contradicted
% by more recent studies of radio spectral indices \citep{dipompeo2012}.
ADWs offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.

Within the ADW  unification framework, the broad-line region (BLR) can 
correspond either to the dense wind base or dense clumps embedded
in the outflows. Indeed, \cite{elitzur2014} show that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit high ionization line absorption; LoBALQSOs, which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs, which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified geometric models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

% As well as acting as a source of photoionized plasma, a 
% As well as imprinting clear line absorption and emission
% features, disc winds may also have a profound effect on the structure and 
% emergent {\em continuum} of the accretion disc itself.
% Mass-loss will alter the accretion rate and resultant 
% temperature of the accretion disc, possibly explaining some 
% of the features typically seen in luminous AGN \citep{knigge1999,laordavis2014}.
% There have been numerous difficulties when confronting 
% theoretical accretion disc models with observations 
% \citep[see e.g.][]{blaes1998}. However, AGN spectral energy distributions (SEDs) can now, 
% in general, be fitted well with accretion disc models
% when the effects of general relativity (GR), Comptonisation
% and mass-loss are included \citep{capellupo2015}. Mass-loss therefore appears to be 
% critical if an accretion disc model is to successfully fit AGN SEDs, 
% particular in the UV region of the spectrum.

%{\bf ksl: I don't see what this paragraph has to do with our problem, which we need to clearly explain} Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics,
Despite the clear importance of ADWs in shaping quasar and AGN spectra,  
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms for ADWs have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
% The presence of line-locked features \citep{bowler2014} 
% and the `ghost of \la' (Arav et al. 1996; Arav 1996; North 2006; but see 
% also Cottis et al. 2010) \nocite{arav1995, arav1996, north2006,cottis2010}
% in the spectra of some BALQSOs also gives clearer evidence that line-driving is
% at least partially contributing to the acceleration of the wind.
The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma, meaning that it is difficult to prevent 
the wind becoming over-ionized and `failing' in the presence of strong X-rays. 
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
% Hydrodynamic simulations of line-driven disc winds also found a shielding region
% was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
% However, \cite{sim2010_hydro} and \cite{H14} showed that including multiple scattering means the ionizing radiation field could still reach the previously shielded regions 
% in those particular models.
An additional or alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \citep{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required to explain the electron scattering wings of emission lines formed
in line-driven hot star winds \citep{hillier1991eswingsmodel}. Complex substructures 
are also produced in simulations of line-driven 
outflows in AGN, although on very different scales to line-driven instabilities 
\citep{PSK2000,PK04,progakurosawa2010,proga2014}.
Nevertheless, clumpy winds offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

@@

Our approach to line transfer is based upon the macro-atom implementation developed by \cite{lucy2002, lucy2003}. In our case, for reasons of computational effieciency, we adopt the  hybrid macro-atom scheme 
described by M15 
%(an improvement over that used by H14 for our earlier quasar study), 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interact with either two-level `simple ions'
or full `macro-atoms'. 
Our scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H is treated as a macro-atom, because 
we expect recombination to be important
in determining their level populations and resultant line emission, and because we were especially interested in the contribution to AGN spectra of \LA.  In H13, all atoms were treated in a two-level approximation.

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by Lucy (2002,2003). Previously (LK02, H13, M15),
we used a modified Saha approach to calculate the ionization fractions
of simple-ions. As part of  this effort, we have now improved {\sc PYTHON} to explicitly solve the 
rate equations between ions in non-LTE. This dispenses with a number of small assumptions 
made in the modified Saha approach, is more numerically stable, and, in principle, allows 
the direct addition of extra physical processes that would previously have necessitated 
approximate treatments.

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{DR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination, respectively.
%% We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

\subsection{Physical Processes}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% MODEL DESCRIPTION
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Our kinematic prescription for a biconical disc wind model
follows \cite{SV93}, and is described further by
LK02, H13 and M15. A schematic is shown in figure~\ref{fig:cartoon},
with key aspects marked. The general biconical
geometry is similar to that invoked by \cite{MCGV95} and 
\cite{elvis2000} to explain the phenomenonology
of quasars and BALQSOs.

@@

@@

The accretion disc in our model is geometrically thin, but optically thick.
We thus treat the disc as an ensemble of blackbodies with a 
\cite{shakurasunyaev1973} effective temperature profile. 
The emergent SED is then determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are assumed to be opaque, meaning
that photons which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~r_G$, where 
$r_G = GM_{BH}/c^2$ is the gravitational radius.
For a $10^9~M_\odot$ black hole, this is equal to $8.8\times10^{14}~{\rm cm}$ 
or $\sim10^{-4}~{\rm pc}$.

The X-ray source in our model is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet
constraint (see Lucy 2002, M15).

@@

In our model, a biconical disc wind rises from the accretion 
disc between launch radii $r_{min}$ and $r_{max}$.
The opening angles of the wind are set to $\theta_{min}$ and $\theta_{max}$.
The poloidal velocity along each individual streamline at a poloidal distance $l$ 
is then given by
\begin{equation}
v_l=v_0+\left[v_{\infty}(r_0)-v_0\right]\frac{\left(l/R_v\right)^{\alpha}}{\left(l/R_v\right)^{\alpha}+1},
\label{v_law}
\end{equation}
where $v_0$ is the velocity at the base of the streamline, $\alpha$ is
an exponent governing how quickly the wind accelerations and 
$R_v$ is the `acceleration length', defined as the distance at which
the outflow reaches half of its terminal velocity, $v_{\infty}$.
The terminal velocity is set to a fixed multiple of the escape
velocity, $v_{esc}$, at the base of the streamline (radius $r_0$).
The rotational velocity, $v_{\phi}$, is initially Keplerian ($v_k = [GM/r_0]^{1/2}$),
and the wind conserves specific angular momentum, such that 
\begin{equation}
v_{\phi} r = v_k r_0.
\label{v_law}
\end{equation}
The velocity law is crucial in determining the output spectra,
as it affects not only the projected velocities along the line of sight,
but also the density and ionization state of the outflow.
A wind that accelerates more slowly will have a denser wind base
with correspondingly different ionization and emission characteristics.

@@

To allow for clumping in our outflow we adopt a simple approximation
used extensively in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998}(MORE REFS). The key assumption here is that 
typical clump sizes are much smaller than the typical photon mean free path, 
and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `filling factor', $f$, which is the 
fraction of the volume of the plasma filled by clumps. The intra-clump
medium is assumed to be a vacuum. We can then introduce the 
density enhancement, $D$, which is defined as 
\begin{equation}
D = \frac{1}{f}.
\end{equation}
We then multiply all densities in the model by $D$, and all emitting volumes
by $f$. This has the effect of enhancing all emissivities and opacities
that scale with the square of density (such as collisional excitation and recombination) 
by a factor $D$. All processes that scale linearly with density 
(such as electron scattering and bound-free opacity)
will remain unchanged for a given ionization state.

@@

\subsection{The Simulation Grid: Arriving at a next-generation model}

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}). 
We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce strong BALs with $BI > ??$?
\item Does significant line emission emerge at low inclinations, with $EW\sim50\AA$ in \civ?
\item Do H recombination lines appear in the spectrum, $EW\sim50\AA$ in \la?
\item Do a certain range of angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
We arrived at a model
In this section, we present one of the most promising models and discuss
the various successes and failures.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

\begin{table}
\begin{tabular}{p{2cm}p{1cm}p{1cm}p{1cm}p{1cm}}
Parameter & \multicolumn{4}{|l|}{Grid Point Values}  \\
\hline \hline 
$r_{min}$ 	&	 $6r_{g}$ & $60r_{g}$ & \multicolumn{2}{|l|}{$300r_{g}$} \\ 
$\theta_{min}$ 	&	 $40^{\circ}$ & $55^{\circ}$ & \multicolumn{2}{|l|}{$70^{\circ}$} \\ 
$R_v$  	        &	 $10^{18}$cm & \multicolumn{3}{|l|}{$10^{19}$cm} \\ 
$\alpha$ 	&	 $0.5$ & $0.6$ & $0.75$ & $1.5$ \\
$f$ 	&	 $0.01$ & $0.1$ & \multicolumn{2}{|l|}{$1$}  \\
\hline 
\end{tabular}
\caption{The grid points used in the parameter search.}
\label{grid_table}
\end{table}

@@

% {\bf We need to describe what we did before getting to results.  We started with parameters similar to those of H14, we created a small grid.  Limits of the grid are given in table 2.  This should be a paragraph or section before this one.  Not all quasars have to have the same parameters and so we have to explain that we have elected to describe one model in detail.  Was our goal to fix the luminosity and see what was required.
